---
title: "FE-582: Financial Data Science  \nProject Proposal"
author: "Daniel Bachalis, Gurjivan Kalkat, Naveen Nagarajan"
date: 'Due: March 16, 2021'
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/Dan/Stevens/Financial Data Science/Project")
library(tidyverse)
library(readxl)
options(scipen=999)
library('gdata')
library('ggplot2')
library('plyr')
library('doBy')
library('readxl') 
library('pastecs')
library(ggpubr)
library(ISLR)
library(splines)
library(gam)
library(tree)
library(randomForest)
library(corrplot)
library(GGally)
library(quantmod)
library(fBasics)
library(xts)
options(scipen=999)
```

\#\#Exploratory Data Analysis (EDA)

\#\#\#Covid and TSA Datasets (Gurj) \* Summarize all of the datasets \*
Load in datasets for TSA and Covid data \* Explore and compare visually,
look for outliers and anything interesting

```{r}
#Load Libraries
library('gdata')
library('ggplot2')
library('plyr')
library('doBy')
library('readxl') 
library('pastecs')

#Load the Libraries
TSA_COVID_BTC <- read_excel("Clean Project Data-Based on Date.xlsx",1)
TLT_DJIA_NDX_GOLD_OIL<-read_excel("Clean Project Data-Based on Date.xlsx",2)
GDP_Uneployment<-read_excel("Clean Project Data-Based on Date.xlsx",3)

#Coverting to Dates
TSA_COVID_BTC$Date<-as.Date(as.character(TSA_COVID_BTC$Date),format = "%Y-%m-%d")
TLT_DJIA_NDX_GOLD_OIL$Date<-as.Date(as.character(TLT_DJIA_NDX_GOLD_OIL$Date),format = "%Y-%m-%d")
GDP_Uneployment$Date<-as.Date(as.character(GDP_Uneployment$Date),format = "%Y-%m-%d")

#Data Stats
Stats1<- TSA_COVID_BTC[-1]
TSA_COVID_BTC_Stats<-stat.desc(Stats1)
format(TSA_COVID_BTC_Stats,scientific=FALSE)


Stats2<- TLT_DJIA_NDX_GOLD_OIL[-1]
TLT_DJIA_NDX_GOLD_OIL_Stats<-stat.desc(Stats2)
format(TLT_DJIA_NDX_GOLD_OIL_Stats,scientific=FALSE)

Stats3<- GDP_Uneployment[-1]
GDP_Uneployment_Stats<-stat.desc(Stats3)
format(GDP_Uneployment_Stats,scientific=FALSE)

#Time Series for TSA and COVID
ggplot(data=TSA_COVID_BTC, aes(x=Date, y=TSA)) +
  geom_line()
ggplot(data=TSA_COVID_BTC, aes(x=Cases, y=Deaths)) +
  geom_line()
ggplot(data=TSA_COVID_BTC, aes(x=Cases, y=Rate)) +
  geom_line()
ggplot(data=TSA_COVID_BTC, aes(x=Cases, y=TSA)) +
  geom_line()
ggplot(data=TSA_COVID_BTC, aes(x=Date, y=Deaths, na.rm=TRUE)) +
  geom_line()
```

\#\#\#Nasdaq, Gold and Oil (Dan) \* Load in datasets for Nasdaq (NDX),
Gold, Oil futures and Covid data \* Explore and compare visually, look
for outliers and anything interesting

```{r}
# Load in datasets
Covid <- read_excel("Clean Project Data-Based on Date.xlsx", 1)
NDX_Gold_Oil <- read_excel("Clean Project Data-Based on Date.xlsx", 2)

# Format the dates
Covid$Date <- as.Date(as.character(Covid$Date), format = "%Y-%m-%d")
NDX_Gold_Oil$Date <- as.Date(as.character(NDX_Gold_Oil$Date), format = "%Y-%m-%d")

# Combine datasets
Covid_NDX_Gold_Oil <- merge(Covid[, c("Date", "Cases", "Rate", "Deaths")], NDX_Gold_Oil[, c("Date", "NDX", "Gold", "Oil")], by = "Date")

# Check the dataset
head(Covid_NDX_Gold_Oil)
tail(Covid_NDX_Gold_Oil)


# Summarize the datasets
summary(Covid_NDX_Gold_Oil)

# Visualize and explore the datasets

# Played around here.  This did not prove very useful all in one plot.
#colors <- c("Deaths" = "red", "Cases" = "green", "NDX" = "blue", "Gold" = "yellow", "Oil" = "black")
#ggplot(Covid_NDX_Gold_Oil, aes(x = Date)) +
#  geom_line(aes(y = NDX, color = "NDX")) +
#  geom_line(aes(y = Gold, color = "Gold")) +
#  geom_line(aes(y = Oil, color = "Oil")) +
#  geom_line(aes(y = Deaths, color = "Deaths")) +
#  geom_line(aes(y = Cases, color = "Cases")) +
#  labs(title="Nasdaq, Gold and Oil with Covid Cases and Deaths", x ="Date", y = "Price", color = "Legend") +
#  scale_color_manual(values = colors)


ggplot(Covid_NDX_Gold_Oil, aes(x = Date, y = Deaths, na.rm = TRUE)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Date, y = NDX)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Date, y = Gold)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Date, y = Oil)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Cases, y = NDX)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Cases, y = Gold)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Cases, y = Oil)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Deaths, y = NDX)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Deaths, y = Gold)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Deaths, y = Oil)) +
  geom_line()
ggplot(Covid_NDX_Gold_Oil, aes(x = Cases, y = Deaths)) +
  geom_line()

```

\#\#\#Dow, XBT and Macroeconomic (Naveen) \* Load in datasets for Dow,
XBT and macro data \* Explore and compare visually, look for outliers
and anything interesting

```{r}
library(ggpubr)
#Combine data set

# Load in datasets
Covid_BTC_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 1)
SPX_DOW_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 2)
MACRO_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 3)

# Format the dates
Covid_BTC_DS$Date <- as.Date(as.character(Covid_BTC_DS$Date), format = "%Y-%m-%d")
SPX_DOW_DS$Date <- as.Date(as.character(SPX_DOW_DS$Date), format = "%Y-%m-%d")
MACRO_DS$Date <- as.Date(as.character(MACRO_DS$Date), format = "%Y-%m-%d")

# Combine datasets
COVID_BTC_SPX_DOW <- right_join(Covid_BTC_DS, SPX_DOW_DS, by = "Date")

COVID_BTC_SPX_DOW_MAC <- left_join(COVID_BTC_SPX_DOW,MACRO_DS,by = "Date")

COVID_BTC_SPX_DOW_MAC <- tibble(COVID_BTC_SPX_DOW_MAC)

# Check the dataset
head(COVID_BTC_SPX_DOW_MAC)
tail(COVID_BTC_SPX_DOW_MAC)


# SPX Plot
plot1<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Date, y = Cases)) +
  geom_line() 
plot2<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Date, y = Deaths)) +
  geom_line() 
plot3<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Date, y = SPX)) +
  geom_line()
plot4<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Date, y = Bitcoin)) +
  geom_line()
plot5<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Date, y = DJIA)) +
  geom_line()
plot6 <- COVID_BTC_SPX_DOW_MAC %>% filter(GDP>0) %>% ggplot(aes(x = Date, y = GDP)) +
  geom_line()
plot7 <- COVID_BTC_SPX_DOW_MAC %>% filter(Unemployment>0) %>% ggplot(aes(x = Date, y = Unemployment)) +
  geom_line()

ggarrange(plot1,plot2,plot3,plot4,plot5,plot6,plot7,nrow = 4, ncol=2)


plot8<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Cases, y = SPX, na.rm = TRUE)) +
  geom_line() 
plot9<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Cases, y = Bitcoin, na.rm = TRUE)) +
  geom_line() 
plot10<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Cases, y = DJIA, na.rm = TRUE)) +
  geom_line() 
plot11<- COVID_BTC_SPX_DOW_MAC %>% filter(GDP>0) %>% ggplot( aes(x = Cases, y = GDP, na.rm = TRUE)) +
  geom_line() 
plot12<- COVID_BTC_SPX_DOW_MAC %>% filter(GDP>0) %>% ggplot( aes(x = Cases, y = Unemployment, na.rm = TRUE)) +
  geom_line() 
ggarrange(plot8,plot9,plot10,plot11,plot12,nrow = 3, ncol=2)

plot13<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Deaths, y = SPX, na.rm = TRUE)) +
  geom_line() 
plot14<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Deaths, y = Bitcoin, na.rm = TRUE)) +
  geom_line() 
plot15<-ggplot(COVID_BTC_SPX_DOW_MAC, aes(x = Deaths, y = DJIA, na.rm = TRUE)) +
  geom_line() 
plot16<- COVID_BTC_SPX_DOW_MAC %>% filter(GDP>0) %>% ggplot( aes(x = Deaths, y = GDP, na.rm = TRUE)) +
  geom_line() 
plot17<- COVID_BTC_SPX_DOW_MAC %>% filter(GDP>0) %>% ggplot( aes(x = Deaths, y = Unemployment, na.rm = TRUE)) +
  geom_line() 
ggarrange(plot13,plot14,plot15,plot16,plot17,nrow = 3, ncol=2)

```

-   Using forward selection, backward selection, and both forward and
    backward selction to find the best model for NDX, Gold and Oil price
-   Evaluate whether this model makes sense and test

```{r}
library(quantmod)

# Build multiple regression model for all three factors of Covid vs. NDX price and summarize
NDX.model <- lm(NDX ~ Deaths + Cases + Rate, data = Covid_NDX_Gold_Oil)
summary(NDX.model)

# Build null model for NDX taste and summarize
NDX.null <- lm(NDX ~ 1, data = Covid_NDX_Gold_Oil)
summary(NDX.null)

# Set up scope for searching model
NDX.model.formula <- NDX ~ Deaths + Cases + Rate

# Evaluate model using forward selection
# step(NDX.null, NDX.model.formula, direction = "forward") # Due to missing/NA values, Forward will have issues.

# Evaluate model using bacward selection
step(NDX.model, NDX.model.formula, direction = "backward")

# Evaluate model using both
# step(NDX.null, NDX.model.formula, direction = "both") # Due to missing/NA values, Forward will have issues.


# Repeat for Gold
Gold.model <- lm(Gold ~ Deaths + Cases + Rate, data = Covid_NDX_Gold_Oil)
Gold.null <- lm(Gold ~ 1, data = Covid_NDX_Gold_Oil)
Gold.model.formula <- Gold ~ Deaths + Cases + Rate
step(Gold.model, Gold.model.formula, direction = "backward")

# Repeat for Oil
Oil.model <- lm(Oil ~ Deaths + Cases + Rate, data = Covid_NDX_Gold_Oil)
Oil.null <- lm(Oil ~ 1, data = Covid_NDX_Gold_Oil)
Oil.model.formula <- Oil ~ Deaths + Cases + Rate
step(Oil.model, Oil.model.formula, direction = "backward")
```

Based on this criteria, the best model for each price are the following:
$$NDX = 8.358e{03} + 1.537e{-02} * Deaths - 8.345e{-05} * Cases + 1.926e{00} * Rate$$
$$Gold = 1.584e{03} + 2.685e{-03} * Deaths - 4.071e{-05} * Cases + 3.586e{-01} * Rate$$
$$Oil = 4.248e{01} - 5.137e{-05} * Deaths + 1.642e{-06} * Cases - 1.836e{-02} * Rate$$
These models probably are not very good.



#### Random Forest and Step function implementation

```{r}


read_excel("Clean Project Data-Based on Date.xlsx", 1)

# Load in datasets
Covid_BTC_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 1)
SPX_DOW_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 2)
MACRO_DS <- read_excel("Clean Project Data-Based on Date.xlsx", 3)

# Format the dates
Covid_BTC_DS$Date <- as.Date(as.character(Covid_BTC_DS$Date), format = "%Y-%m-%d")
SPX_DOW_DS$Date <- as.Date(as.character(SPX_DOW_DS$Date), format = "%Y-%m-%d")
MACRO_DS<- rbind(MACRO_DS,c('2021-03-02',0,0,0))
MACRO_DS$Date <- as.Date(as.character(MACRO_DS$Date), format = "%Y-%m-%d")
MACRO_DS$GDP<-as.numeric(MACRO_DS$GDP)
MACRO_DS$Unemployment<-as.numeric(MACRO_DS$Unemployment)

MACRO_DS<- MACRO_DS %>%
  complete(Date = seq.Date(min(Date), max(Date), by="day")) %>% fill('GDP') %>% fill('Unemployment') 


# Combine datasets
covid.full.ds <- right_join(Covid_BTC_DS, SPX_DOW_DS, by = "Date")
covid.full.ds <- left_join(covid.full.ds,MACRO_DS,by = "Date")

# replacing NA with 0
covid.full.ds<- covid.full.ds %>%
  mutate(Cases = if_else(is.na(Cases), 0, Cases)) %>%
  mutate(Deaths = if_else(is.na(Deaths), 0, Deaths))

summary(covid.full.ds)

# Check the dataset
head(covid.full.ds)
tail(covid.full.ds)

ggcorr(covid.full.ds %>% dplyr::select(Date:Cases,Deaths:Unemployment), method = c("everything", "pearson"))


covid.full.ds<- covid.full.ds %>% dplyr::mutate(rnum = row_number())

# calculate retursn and diff
covid.full.ds<-data.frame(covid.full.ds %>% arrange(Date) %>% dplyr::select(Date:rnum),
                          c(0,diff(covid.full.ds$Cases)),
                          c(0,diff(covid.full.ds$Deaths)),
                          c(0,diff(log(covid.full.ds$SPX))),
                          c(0,diff(log(covid.full.ds$Bitcoin))),
                          c(0,diff(log(covid.full.ds$DJIA))),
                          c(0,diff(log(covid.full.ds$NDX))),
                          c(0,diff(log(covid.full.ds$TLT))),
                          c(0,diff(log(covid.full.ds$Gold))),
                          c(0,diff(log(covid.full.ds$Oil)))
) %>% arrange(Date)
names(covid.full.ds)[16]<-"Cases.diff"
names(covid.full.ds)[17]<-"Deaths.diff"
names(covid.full.ds)[18]<-"SPX.ret"
names(covid.full.ds)[19]<-"Bitcoin.ret"
names(covid.full.ds)[20]<-"DJIA.ret"
names(covid.full.ds)[21]<-"NDX.ret"
names(covid.full.ds)[22]<-"TLT.ret"
names(covid.full.ds)[23]<-"Gold.ret"
names(covid.full.ds)[24]<-"Oil.ret"


#Split into test and train
covid.full.ds.train <- covid.full.ds %>% dplyr::filter(Date<as.Date("2021-01-10"))
covid.full.ds.test  <- anti_join(covid.full.ds, covid.full.ds.train, by = 'rnum')

covid.full.ds.train.1<- covid.full.ds.train %>% dplyr::select(Date:Cases,Deaths:Unemployment,Cases.diff:Deaths.diff)
covid.full.ds.test.1<- covid.full.ds.test %>% dplyr::select(Date:Cases,Deaths:Unemployment,Cases.diff:Deaths.diff)


covid.full.ds.train.2<- covid.full.ds.train %>% dplyr::select(Date,Cases.diff:Deaths.diff,GDP,Unemployment,SPX.ret:Oil.ret) %>% dplyr::filter(Date>as.Date("2019-01-02"))
covid.full.ds.test.2<- covid.full.ds.test %>% dplyr::select(Date,Cases.diff:Deaths.diff,GDP,Unemployment,SPX.ret:Oil.ret) 

pairs(covid.full.ds.train %>% dplyr::select(Date:Cases,Deaths:Unemployment))
#cor(covid.full.ds.train %>% dplyr::select(Date:Cases,Deaths:Unemployment))

#corrplot(covid.full.ds.train.1 , method="number", use="complete.obs", is.corr=F)

tail(covid.full.ds.test)

#RandomForest

##SPX

set.seed(3)
rf.covid.spx<-randomForest(SPX~.,data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,-c(NDX,DJIA,Bitcoin)), 
                       mtry=4, importance=TRUE)
yhat.rf.spx<-predict(rf.covid.spx, newdata = covid.full.ds.test.1)
mean((yhat.rf.spx-covid.full.ds.test.1$SPX)^2)
importance(rf.covid.spx)
varImpPlot(rf.covid.spx)

##Bitcoin
rf.covid.bitcoin<-randomForest(Bitcoin~.,data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,-c(NDX,DJIA,SPX)), 
                           mtry=4, importance=TRUE)
yhat.rf.bitcoin<-predict(rf.covid.bitcoin, newdata = covid.full.ds.test.1)
mean((yhat.rf.bitcoin-covid.full.ds.test.1$Bitcoin)^2)
importance(rf.covid.bitcoin)
varImpPlot(rf.covid.bitcoin)

#DJIA
rf.covid.djia<-randomForest(DJIA~.,data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,-c(NDX,Bitcoin,SPX)), 
                               mtry=4, importance=TRUE)
yhat.rf.djia<-predict(rf.covid.djia, newdata = covid.full.ds.test.1)
mean((yhat.rf.djia-covid.full.ds.test.1$DJIA)^2)
importance(rf.covid.djia)
varImpPlot(rf.covid.djia)


#NDX
rf.covid.ndx<-randomForest(NDX~.,data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,-c(DJIA,Bitcoin,SPX)), 
                            mtry=4, importance=TRUE)
yhat.rf.ndx<-predict(rf.covid.ndx, newdata = covid.full.ds.test.1)
mean((yhat.rf.ndx-covid.full.ds.test.1$NDX)^2)
importance(rf.covid.ndx)
varImpPlot(rf.covid.ndx)


#Step Function

#SPX

step.covid.fit.spx<- gam(SPX~cut(Cases.diff,breaks=c(-1,1,10000,30000,70000,400000,900000))+
                           cut(Deaths.diff,breaks=c(-1,1,1000,3000,5000,10000))+TSA+TLT+Unemployment+GDP+Gold+Oil, 
                    data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,Cases.diff:Deaths.diff,-c(NDX,DJIA,Bitcoin)))
yhat.step.covid.spx<-predict(step.covid.fit.spx, newdata = covid.full.ds.test.1)
mean((yhat.step.covid.spx-covid.full.ds.test.1$SPX)^2)
par(mfrow=c(1,1))
plot(yhat.step.covid.spx, se=TRUE,col="blue")

#Bitcoin
step.covid.fit.btc<- gam(Bitcoin~cut(Cases.diff,breaks=c(-1,1,10000,30000,70000,400000,900000))+
                           cut(Deaths.diff,breaks=c(-1,1,1000,3000,5000,10000))+TSA+TLT+Unemployment+GDP+Gold+Oil, 
                         data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,Cases.diff:Deaths.diff,-c(NDX,DJIA,SPX)))
yhat.step.covid.btc<-predict(step.covid.fit.btc, newdata = covid.full.ds.test.1)
mean((yhat.step.covid.btc-covid.full.ds.test.1$Bitcoin)^2)
par(mfrow=c(1,1))
plot(yhat.step.covid.btc, se=TRUE,col="blue")

#DJIA
step.covid.fit.djia<- gam(DJIA~cut(Cases.diff,breaks=c(-1,1,10000,30000,70000,400000,900000))+
                           cut(Deaths.diff,breaks=c(-1,1,1000,3000,5000,10000))+TSA+TLT+Unemployment+GDP+Gold+Oil, 
                         data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,Cases.diff:Deaths.diff,-c(NDX,Bitcoin,SPX)))
yhat.step.covid.djia<-predict(step.covid.fit.djia, newdata = covid.full.ds.test.1)
mean((yhat.step.covid.djia-covid.full.ds.test.1$DJIA)^2)
par(mfrow=c(1,1))
plot(yhat.step.covid.djia, se=TRUE,col="blue")

#NDX
step.covid.fit.ndx<- gam(NDX~cut(Cases.diff,breaks=c(-1,1,10000,30000,70000,400000,900000))+
                            cut(Deaths.diff,breaks=c(-1,1,1000,3000,5000,10000))+TSA+TLT+Unemployment+GDP+Gold+Oil, 
                          data=covid.full.ds.train.1 %>% dplyr::select(TSA:Unemployment,Cases.diff:Deaths.diff,-c(DJIA,Bitcoin,SPX)))
yhat.step.covid.ndx<-predict(step.covid.fit.ndx, newdata = covid.full.ds.test.1)
mean((yhat.step.covid.ndx-covid.full.ds.test.1$NDX)^2)
par(mfrow=c(1,1))
plot(yhat.step.covid.djia, se=TRUE,col="blue")

```
